\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2016-01-12 nnt_manu.tex
everything sweave, no more knitr

\documentclass[article,shortnames]{jss}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

% knitr options

% get online bib file

%% need no \usepackage{Sweave.sty}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

Methods in \pkg{NeuralNetTools} were written for several \proglang{R} packages that can be used to create \ac{mlp} neural networks: \pkg{neuralnet} \citep{Fritsch12}, \pkg{nnet} \citep{Venables02}, and \pkg{RSNNS} \citep{Bergmeir12}. Limited methods were also developed for neural network objects created with the \code{train} function from the \pkg{caret} package \citep{Kuhn15}.  Additional \proglang{R} packages that can create \ac{mlp} neural networks include \pkg{AMORE} that implements the `TAO-robust back-propagation algorithm' for model fitting \citep{Castejon14}, \pkg{FCNN4R} as an \proglang{R} interface to the \pkg{FCNN} \proglang{C++} library \citep{Klima15}, \pkg{monmlp} for networks with partial monotonicity constraints \citep{Cannon15}, and \pkg{qrnn} for quantile regression neural networks \citep{Cannon11}.  At the time of writing, the \ac{CRAN} download logs \citep{Csardi15} showed that the \proglang{R} packages with methods in \pkg{NeuralNetTools} included 94\% of all downloads for the available \ac{mlp} packages, with \pkg{nnet} accounting for over 79\%.  As such, methods have not been included in \pkg{NeuralNetTools} for the remaining packages, although further development of \pkg{NeuralNetTools} could include additional methods based on popularity.  Methods for each function are currently available for  \code{mlp} (\pkg{RSNNS}), \code{nn} (\pkg{neuralnet}), \code{nnet} (\pkg{nnet}), and \code{train} (\pkg{caret}, only if the object also inherits from the \code{nnet} class.  Additional \code{default} or \code{numeric} methods are available for some of the generic functions.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # install from CRAN
R> install.packages('NeuralNetTools')
R> library(NeuralNetTools)
R> 
R> # install from GitHub
R> install.package('devtools')
R> library(devtools)
R> install_github('fawda123/NeuralNetTools')
R> library(NeuralNetTools)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # set random seed for initial weights for training
R> set.seed(123)
R> 
R> # mlp object, RSNNS package
R> library(RSNNS)
R> x <- neuraldat[, c('X1', 'X2', 'X3')]
R> y <- neuraldat[, 'Y1']
R> mod1 <- mlp(x, y, size = 5)
R> 
R> # nn object, neuralnet package
R> library(neuralnet)
R> mod2 <- neuralnet(Y1 ~ X1 + X2 + X3, data = neuraldat, hidden = 5)
R> 
R> #nnet object, nnet package
R> library(nnet)
R> mod3 <- nnet(Y1 ~ X1 + X2 + X3, data = neuraldat, size = 5)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> par(mar = c(0, 0, 0, 0))
R> plotnet(mod3, nid = F, circle_col = 'grey', bord_col = 'grey')
R> plotnet(mod3)
\end{verbatim}
\end{kframe}\begin{figure}[!ht]

{\centering \subfloat[\label{fig:plotnet1}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-1} }
\subfloat[\label{fig:plotnet2}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-2} }

}

\caption{Examples from the \code{plotnet} function showing neural networks as a standard graphic (\ref{fig:plotnet1}) and using the \acl{nid} (\ref{fig:plotnet2}).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).}\label{fig:plotnet}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # create a model with a skip layer
R> modskip <- nnet(Y1 ~ X1 + X2 + X3, data = neuraldat, size = 5, skip = TRUE)
R> 
R> # plot
R> par(mar = c(0, 0, 0, 0))
R> plotnet(modskip, skip = TRUE)
R> plotnet(modskip)
\end{verbatim}
\end{kframe}\begin{figure}[!ht]

{\centering \subfloat[\label{fig:plotnet_skip1}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-1} }
\subfloat[\label{fig:plotnet_skip2}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-2} }

}

\caption{Examples from the \code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (\ref{fig:plotnet_skip1}) and primary neural network (\ref{fig:plotnet_skip2}) can be viewed separately with \code{plotnet} by using \code{skip = TRUE} or \code{skip = FALSE}.}\label{fig:plotnet_skip}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # pruned model using RSNNS
R> pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
+   no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, 
+   init_matrix_value = 1e-6, input_pruning = TRUE, hidden_pruning = TRUE)
R> mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
+  pruneFuncParams = pruneFuncParams)
R> 
R> # plot, pruned connections are omitted (default) or included with prune_col
R> par(mar = c(0, 0, 0, 0))
R> plotnet(mod, rel_rsc = c(3, 8))
R> plotnet(mod, prune_col = 'lightblue', rel_rsc = c(3, 8))
\end{verbatim}
\end{kframe}\begin{figure}[!ht]

{\centering \subfloat[\label{fig:plotprune1}]{\includegraphics[width=0.48\textwidth]{figs/plotprune-1} }
\subfloat[\label{fig:plotprune2}]{\includegraphics[width=0.48\textwidth]{figs/plotprune-2} }

}

\caption{A pruned neural network from \pkg{RSNNS} \citep{Bergmeir12} using the `Optimal Brain Surgeon' algorithm described in \citet{Zell98}.  The default plotting behavior of \code{plotnet} is to omit pruned connections (\ref{fig:plotprune1}), whereas they can be viewed as dashed lines by including the \code{prune\_col} argument (\ref{fig:plotprune2}).}\label{fig:plotprune}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # garson and olden functions for each model
R> garson(mod1)
R> olden(mod1)
R> garson(mod2)
R> olden(mod2)
R> garson(mod3)
R> olden(mod3)
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
\centering
\subfloat[][Model 1 (\code{mlp}), \code{garson}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-1-1}

}

\end{knitrout}
\label{fig:plotimp1}
}
\subfloat[][Model 1 (\code{mlp}), \code{olden}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-2-1}

}

\end{knitrout}
\label{fig:plotimp2}
}

\subfloat[][Model 2 (\code{nn}), \code{garson}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-3-1}

}

\end{knitrout}
\label{fig:plotimp3}
}
\subfloat[][Model 2 (\code{nn}), \code{olden}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-4-1}

}

\end{knitrout}
\label{fig:plotimp4}
}

\subfloat[][Model 3 (\code{nnet}), \code{garson}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-5-1}

}

\end{knitrout}
\label{fig:plotimp5}
}
\subfloat[][Model 3 (\code{nnet}), \code{olden}]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-6-1}

}

\end{knitrout}
\label{fig:plotimp6}
}
\caption{Variable importance for three models using Garson's algorithm for relative importance (\code{garson}, \cref{fig:plotimp1,fig:plotimp3,fig:plotimp5}, \citealt{Garson91,Goh95}) and Olden's conection weights algorithm (\code{olden}, \cref{fig:plotimp2,fig:plotimp4,fig:plotimp6}, \citealt{Olden04}).  Garson's algorithm shows importance as absolute values from 0--1, whereas Olden's algorithm preserves sign and magnitude.  Importance values for Olden's algorithm are from the summed product of model weights and are not rescaled.}
\label{fig:plotimp}
\end{figure}

@@

A second implementation of \code{lekprofile} is to group the unevaluated explanatory variables by natural groupings defined by the data. Covariance among predictors may present unlikely scenarios if all unevaluated variables are held at the same level (e.g., high values for one variable may be unlikely with high values for a second variable). The second option holds unevaluated variables at mean values defined by natural clusters in the data (\cref{fig:plotlek_sens2,fig:plotlek_bars2}). Clusters are identified using {\it k}-means clustering (\code{kmeans} from base \pkg{stats}, \citealt{Hartigan79}) of the input variables if the argument passed to \code{group_vals} is an integer value greater than one. The centers (means) of the clusters are then used as constant values for the unevaluated variables.  \citet{Beck14a} provide an example of the clustering method for \code{lekprofile} by evaluating response of a lake health index to different explanatory variables.  Lake clusters were identified given covariance among variables, such that holding explanatory variables at values defined by clusters created more ecologically valid response curves.  Both methods return similar plots, with additional options to visualize the groupings for unevaluated explanatory variables (\cref{fig:plotlek_bars}).  For the latter case, \code{split\_show = TRUE} will return a stacked bar plot for each group with heights within each bar proportional to the constant values.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> # lek profile, quantile evaluation
R> lekprofile(mod3)
R> 
R> # lek profile, quantile groups 
R> lekprofile(mod3, group_show = TRUE)
R> 
R> # lek profile, cluster evaluation
R> lekprofile(mod3, group_vals = 6)
R> 
R> # lek profile, cluster groups
R> lekprofile(mod3, group_vals = 6, group_show = TRUE)
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{figure}
\centering
\subfloat[][Response with quantiles]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.9\textwidth]{figs/plotlek_sens1-1}

}

\end{knitrout}
\label{fig:plotlek_sens1}
}

\subfloat[][Response with clusters]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.9\textwidth]{figs/plotlek_sens2-1}

}

\end{knitrout}
\label{fig:plotlek_sens2}
}s
\caption{Sensitivity analysis of a neural network using the Lek profile method to evaluate the effects of explanatory variables.  \Cref{fig:plotlek_sens1} groups unevaluated explanatory variables at quantiles (minimum, 20\textsuperscript{th}, 40\textsuperscript{th}, 60\textsuperscript{th}, 80\textsuperscript{th}, and maximum percentiles) and \cref{fig:plotlek_sens2} groups by cluster means (six groups).  Values at which explanatory variables are held constant for each group are shown in \cref{fig:plotlek_bars1,fig:plotlek_bars2}.}
\label{fig:plotlek_sens}
\end{figure}

%%%%
\begin{figure}
\centering
\subfloat[][Quantile groupings]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.9\textwidth]{figs/plotlek_bars1-1}

}

\end{knitrout}
\label{fig:plotlek_bars1}
}

\subfloat[][Cluster groupings]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.9\textwidth]{figs/plotlek_bars2-1}

}

\end{knitrout}
\label{fig:plotlek_bars2}
}
\caption{Bar plots for values of unevaluated explanatory variables in each group in \cref{fig:plotlek_sens1,fig:plotlek_sens2}.  \Cref{fig:plotlek_bars1} shows default quantile groupings set at the minimum, 20\textsuperscript{th}, 40\textsuperscript{th}, 60\textsuperscript{th}, 80\textsuperscript{th}, and maximum percentiles.  For example, variables are held at negative values for group 1 (i.e., stacked bars with negative heights) for the minimum value, whereas group 6 holds variables at their maximum (largest positive heights). \Cref{fig:plotlek_bars2} shows the cluster centers for each variable in each group.  Groups in \cref{fig:plotlek_bars2} are random because the input variables are from a standard normal distribution.}
\label{fig:plotlek_bars}
\end{figure}

@@

Several conclusions can be made from \cref{fig:flightimp}.  First, delays in arrival time are negatively related to distance and positively related to departure delays and travel air.  That is, flights arrived later than their scheduled time if flight time was long or if their departure was delayed, whereas flights arrived earlier than scheduled for longer distances.  No conclusions can be made for the other variables because the bounds of uncertainty included zero.  Second, the range of importance estimates varied between the models (i.e., one node  varied +/- 1 and the others +/- 3).  This suggests that the relative importance estimates only have relevance within each model, whereas only the rankings (e.g., least, most important) can be compared between models.  Third and most important, the level of uncertainty for specific variables can be large between model fits for the same architecture.  This suggests that a single model can provide misleading information such that several models may be required to improve the certainty of conclusions.  Additional considerations described above (e.g., criteria for stopping training, use of training and test datasets) can also affect the interpration of model information and should be equally considered during model development.

\begin{figure}[!h]
\centering
\subfloat[][Networks with one node]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.95\textwidth]{figs/flightimp-1-1}

}

\end{knitrout}
\label{fig:flightimp1}
}

\subfloat[][Networks with five nodes]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.95\textwidth]{figs/flightimp-2-1}

}

\end{knitrout}
\label{fig:flightimp2}
}

\subfloat[][Networks with ten nodes]{
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}

{\centering \includegraphics[width=0.95\textwidth]{figs/flightimp-3-1}

}

\end{knitrout}
\label{fig:flightimp3}
}
\caption{Uncertainty in variable importance estimates for three neural networks to evaluate factors related to arrival delays for flights departing New York City.  Three model types with one, five, and ten nodes were evaluated with 100 models with different starting weights for each type.}
\label{fig:flightimp}
\end{figure}

>>>>

\documentclass[article,shortnames]{jss}

@@

% options and libs
% get online bib file

%% need no \usepackage{Sweave.sty}

\begin{document}

@@

Methods in \pkg{NeuralNetTools} were written for several \proglang{R} packages that can be used to create \ac{mlp} neural networks: \pkg{neuralnet} \citep{Fritsch12}, \pkg{nnet} \citep{Venables02}, and \pkg{RSNNS} \citep{Bergmeir12}. Limited methods were also developed for neural network objects created with the \code{train} function from the \pkg{caret} package \citep{Kuhn15}.  Additional \proglang{R} packages that can create \ac{mlp} neural networks include \pkg{AMORE} that implements the `TAO-robust back-propagation algorithm' for model fitting \citep{Castejon14}, \pkg{FCNN4R} as an \proglang{R} interface to the \pkg{FCNN} \proglang{C++} library \citep{Klima15}, \pkg{monmlp} for networks with partial monotonicity constraints \citep{Cannon15}, and \pkg{qrnn} for quantile regression neural networks \citep{Cannon11}.  At the time of writing, the \ac{CRAN} download logs \citep{Csardi15} showed that the \proglang{R} packages with methods in \pkg{NeuralNetTools} included 94\% of all downloads for the available \ac{mlp} packages, with \pkg{nnet} accounting for over 80\%.  As such, methods have not been included in \pkg{NeuralNetTools} for the remaining packages, although further development of \pkg{NeuralNetTools} could include additional methods based on popularity.  Methods for each function are currently available for  \code{mlp} (\pkg{RSNNS}), \code{nn} (\pkg{neuralnet}), \code{nnet} (\pkg{nnet}), and \code{train} (\pkg{caret}, only if the object also inherits from the \code{nnet} class.  Additional \code{default} or \code{numeric} methods are available for some of the generic functions.

@@

\begin{Schunk}
\begin{Sinput}
R> # install from CRAN
R> install.packages('NeuralNetTools')
R> library(NeuralNetTools)
R> # install from GitHub
R> install.package('devtools')
R> library(devtools)
R> install_github('fawda123/NeuralNetTools')
R> library(NeuralNetTools)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # set random seed for initial weights for training
R> set.seed(123)
R> # mlp object, RSNNS package
R> library(RSNNS)
R> x <- neuraldat[, c('X1', 'X2', 'X3')]
R> y <- neuraldat[, 'Y1']
R> mod1 <- mlp(x, y, size = 5)
R> # nn object, neuralnet package
R> library(neuralnet)
R> mod2 <- neuralnet(Y1 ~ X1 + X2 + X3, data = neuraldat, hidden = 5)
R> #nnet object, nnet package
R> library(nnet)
R> mod3 <- nnet(Y1 ~ X1 + X2 + X3, data = neuraldat, size = 5)
\end{Sinput}
\end{Schunk}

@@

\begin{figure}[!ht]
\subfloat[][]{
\includegraphics[width=7.5cm,page=1]{figs/plotnet.pdf}
\label{fig:plotnet1}
}
\subfloat[][]{
\includegraphics[width=7.5cm, page=2]{figs/plotnet.pdf}
\label{fig:plotnet2}
}
\caption{Examples from the \code{plotnet} function showing neural networks as a standard graphic (\ref{fig:plotnet1}) and using the \acl{nid} (\ref{fig:plotnet2}).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).}
\label{fig:plotnet}
\end{figure}

@@

\begin{figure}[!ht]
\subfloat[][]{
\includegraphics[width=7.5cm,page=1]{figs/plotnet_skip.pdf}
\label{fig:plotnet_skip1}
}
\subfloat[][]{
\includegraphics[width=7.5cm, page=2]{figs/plotnet_skip.pdf}
\label{fig:plotnet_skip2}
}
\caption{Examples from the \code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (\ref{fig:plotnet_skip1}) and primary neural network (\ref{fig:plotnet_skip2}) can be viewed separately with \code{plotnet} by using \code{skip = TRUE} or \code{skip = FALSE}.}
\label{fig:plotnet_skip}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
R> # pruned model using RSNNS
R> pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
+   no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, 
+   init_matrix_value = 1e-6, input_pruning = TRUE, hidden_pruning = TRUE)
R> mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
+  pruneFuncParams = pruneFuncParams)
R> # plot, pruned connections are omitted (default) or included with prune_col
R> par(mar = c(0, 0, 0, 0))
R> plotnet(mod, rel_rsc = c(3, 8))
R> plotnet(mod, prune_col = 'lightblue', rel_rsc = c(3, 8))
\end{Sinput}
\end{Schunk}
\begin{figure}[!ht]
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=1]{figs/plotprune.pdf}
\label{fig:plotprune1}
}
\subfloat[][]{
\includegraphics[width=0.48\textwidth, page=2]{figs/plotprune.pdf}
\label{fig:plotprune2}
}
\caption{A pruned neural network from \pkg{RSNNS} \citep{Bergmeir12} using the `Optimal Brain Surgeon' algorithm described in \citet{Zell98}.  The default plotting behavior of \code{plotnet} is to omit pruned connections (\ref{fig:plotprune1}), whereas they can be viewed as dashed lines by including the \code{prune\_col} argument (\ref{fig:plotprune2}).}
\label{fig:plotprune}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
R> # garson and olden functions for each model
R> garson(mod1)
R> olden(mod1)
R> garson(mod2)
R> olden(mod2)
R> garson(mod3)
R> olden(mod3)
\end{Sinput}
\end{Schunk}
\begin{figure}[!h]
\centering
\subfloat[][Model 1 (\code{mlp}), \code{garson}]{
\includegraphics[width = 0.47\textwidth, page = 1]{figs/plotimp.pdf}
\label{fig:plotimp1}
}
\subfloat[][Model 1 (\code{mlp}), \code{olden}]{
\includegraphics[width = 0.47\textwidth, page = 2]{figs/plotimp.pdf}
\label{fig:plotimp2}
}

\subfloat[][Model 2 (\code{nn}), \code{garson}]{
\includegraphics[width = 0.47\textwidth, page = 3]{figs/plotimp.pdf}
\label{fig:plotimp3}
}
\subfloat[][Model 2 (\code{nn}), \code{olden}]{
\includegraphics[width = 0.47\textwidth, page = 4]{figs/plotimp.pdf}
\label{fig:plotimp4}
}

\subfloat[][Model 3 (\code{nnet}), \code{garson}]{
\includegraphics[width = 0.47\textwidth, page = 5]{figs/plotimp.pdf}
\label{fig:plotimp5}
}
\subfloat[][Model 3 (\code{nnet}), \code{olden}]{
\includegraphics[width = 0.47\textwidth, page = 6]{figs/plotimp.pdf}
\label{fig:plotimp6}
}
\caption{Variable importance for three models using Garson's algorithm for relative importance (\code{garson}, \cref{fig:plotimp1,fig:plotimp3,fig:plotimp5}, \citealt{Garson91,Goh95}) and Olden's conection weights algorithm (\code{olden}, \cref{fig:plotimp2,fig:plotimp4,fig:plotimp6}, \citealt{Olden04}).  Garson's algorithm shows importance as absolute values from 0--1, whereas Olden's algorithm preserves sign and magnitude.  Importance values for Olden's algorithm are from the summed product of model weights and are not rescaled.}
\label{fig:plotimp}
\end{figure}

@@

A second implementation of \code{lekprofile} is to group the unevaluated explanatory variables by natural groupings defined by the data. Covariance among predictors may present unlikely scenarios if all unevaluated variables are held at the same level (e.g., high values for one variable may be unlikely with high values for a second variable). The second option holds unevaluated variables at mean values defined by natural clusters in the data (\cref{fig:plotlek_sens2,fig:plotlek_bars2}). Clusters are identified using {\it k}-means clustering (\code{kmeans} from base \pkg{stats}, \citealt{Hartigan79}) of the input variables if the argument passed to \code{group_vals} is an integer value greater than one. The centers (means) of the clusters are then used as constant values for the unevaluated variables.  \citet{Beck14a} provide an example of the clustering method for \code{lekprofile} by evaluating response of a lake health index to different explanatory variables.  Lake clusters were identified given covariance among variables, such that holding explanatory variables at values defined by clusters created more ecologically valid response curves.  Both methods return similar plots, with additional options to visualize the groupings for unevaluated explanatory variables (\cref{fig:plotlek_bars}).  For the latter case, \code{split\_show = TRUE} will return a stacked bar plot for each group with heights within each bar proportional to the constant values.
\begin{Schunk}
\begin{Sinput}
R> # lek profile, quantile evaluation
R> lekprofile(mod3)
R> # lek profile, quantile groups 
R> lekprofile(mod3, group_show = TRUE)
R> # lek profile, cluster evaluation
R> lekprofile(mod3, group_vals = 6)
R> # lek profile, cluster groups
R> lekprofile(mod3, group_vals = 6, group_show = TRUE)
\end{Sinput}
\end{Schunk}
\begin{figure}[!ht]
\centering
\subfloat[][Response with quantiles]{
\includegraphics[width = 0.9\textwidth, page = 1]{figs/plotlek_sens.pdf}
\label{fig:plotlek_sens1}
}

\subfloat[][Response with clusters]{
\includegraphics[width = 0.9\textwidth, page = 2]{figs/plotlek_sens.pdf}
\label{fig:plotlek_sens2}
}
\caption{Sensitivity analysis of a neural network using the Lek profile method to evaluate the effects of explanatory variables.  \Cref{fig:plotlek_sens1} groups unevaluated explanatory variables at quantiles (minimum, 20\textsuperscript{th}, 40\textsuperscript{th}, 60\textsuperscript{th}, 80\textsuperscript{th}, and maximum percentiles) and \cref{fig:plotlek_sens2} groups by cluster means (six groups).  Values at which explanatory variables are held constant for each group are shown in \cref{fig:plotlek_bars1,fig:plotlek_bars2}.}
\label{fig:plotlek_sens}
\end{figure}

%%%%
\begin{figure}
\centering
\subfloat[][Quantile groupings]{
\includegraphics[width = 0.9\textwidth, page = 1]{figs/plotlek_bars.pdf}
\label{fig:plotlek_bars1}
}

\subfloat[][Cluster groupings]{
\includegraphics[width = 0.9\textwidth, page = 2]{figs/plotlek_bars.pdf}
\label{fig:plotlek_bars2}
}
\caption{Bar plots for values of unevaluated explanatory variables in each group in \cref{fig:plotlek_sens1,fig:plotlek_sens2}.  \Cref{fig:plotlek_bars1} shows default quantile groupings set at the minimum, 20\textsuperscript{th}, 40\textsuperscript{th}, 60\textsuperscript{th}, 80\textsuperscript{th}, and maximum percentiles.  For example, variables are held at negative values for group 1 (i.e., stacked bars with negative heights) for the minimum value, whereas group 6 holds variables at their maximum (largest positive heights). \Cref{fig:plotlek_bars2} shows the cluster centers for each variable in each group.  Groups in \cref{fig:plotlek_bars2} are random because the input variables are from a standard normal distribution.}
\label{fig:plotlek_bars}
\end{figure}

@@

Several conclusions can be made from \cref{fig:flightimp}.  First, delays in arrival time are negatively related to distance and positively related to departure delays and travel air.  That is, flights arrived later than their scheduled time if flight time was long or if their departure was delayed, whereas flights arrived earlier than scheduled for longer distances.  No conclusions can be made for the other variables because the bounds of uncertainty included zero.  Second, the range of importance estimates varied between the models (i.e., one node  varied +/- 1 and the others +/- 3).  This suggests that the relative importance estimates only have relevance within each model, whereas only the rankings (e.g., least, most important) can be compared between models.  Third and most important, the level of uncertainty for specific variables can be large between model fits for the same architecture.  This suggests that a single model can provide misleading information such that several models may be required to improve the certainty of conclusions.  Additional considerations described above (e.g., criteria for stopping training, use of training and test datasets) can also affect the interpration of model information and should be equally considered during model development.
\begin{figure}[!h]
\centering
\subfloat[][Networks with one node]{
\includegraphics[width = 0.95\textwidth, page = 1]{figs/flightimp.pdf}
\label{fig:flightimp1}
}

\subfloat[][Networks with five nodes]{
\includegraphics[width = 0.95\textwidth, page = 2]{figs/flightimp.pdf}
\label{fig:flightimp2}
}

\subfloat[][Networks with ten nodes]{
\includegraphics[width = 0.95\textwidth, page = 3]{figs/flightimp.pdf}
\label{fig:flightimp3}
}
\caption{Uncertainty in variable importance estimates for three neural networks to evaluate factors related to arrival delays for flights departing New York City.  Three model types with one, five, and ten nodes were evaluated with 100 models with different starting weights for each type.}
\label{fig:flightimp}
\end{figure}
